Diversification of Australian blind snakes

### 2022-11-30 manuscript.Rmd
minor updates

## Reconstruction of Australian paleo climate
<!-- Consult Alex Skeels -->
To estimate paleoclimate, we used three paleographic digital elevation models (DEMs) from @scoteseAtlas2021, @straumeGlobal2020, and [VALDES]. Each DEM provide paleotopography at resolution of 1&deg; x 1&deg; and 5 myr resolution. We projected the paleoenvironmental reconstructions into an equal-area Behrmann projection of 110km x 110km following methods described in @hagenGen3sis2021. From the projected data we extracted temperature and aridity values across all grid cells that fall within Sahul by manually drawing a polygon that encompassed the continent from 40 mya to the present-day. Because reptiles are sensitive to extremely low and high temperature, we decided to to summarise the reconstructed paleo temperature data into three different data sets: minimum temperature, maximum temperature, and mean temperature.

@@

## Biogeographic history and biome transitions

<!-- ## values and info from 04_BioGeoBEARS_bsm_plots.R -->
We found that _Anilios_ species are mostly found solely or partially in the tropical grassland biome (n = 20) followed by the arid biome (n = 13) and mediterranean biome (n = 10). Closely related species tend to share the same biome (\@ref(fig:fig-biogeobears). 
We identified the DEC+_j_+_x_ model as the most preferred biogeographical history model (delta AICc = -4.47, see Table-supp-BioGeoBEARS). 
Likelihood ratio test (LRT) indicates that both founder events (+*j* parameter) and distance between biome (+x parameter) together provides statistically significant improvement over the simpler DEC model, DEC+x  and DEC+j (X^2^, DF = 2, D-statistic = 26.7, p-value < 0.001).

Based on 50 BioGeoBEARS biogeographic stochastic mapping of our best model, we inferred the tropical grassland biome to be the ancestral biome of Australian _Anilios_ (\@ref(fig:fig-biogeobears)), but the origin of crown *Anilios* is ambiguous. We estimated an average of 71.2 ± 1.3 (mean ± SD) total anagenetic and cladogenetic events occured throughout the phylogeny, out of which the majority were a combination of anagenetic dispersal (31.2 ± 1.3 events) and founder events (6.1 ± 1.5 events) (see supp-summary_counts_bsm_table). 
We found that the largest source of outward dispersal is the tropical grassland biome (16.2 ± 4.9 events; 43.6%) making up almost double the number of outward dispersals from the arid biome (9.7 ± 3.2 events; 26.0%) (\@ref(fig:fig-biogeobears))c-d. Together the tropical grassland biome and the arid biome make up the majority (69.6%) of outward transitions.

@@

>>>>

## Reconstruction of Australian paleo climate
<!-- Consult Alex Skeels -->
To estimate paleoclimate, we used three paleographic digital elevation models (DEMs) from @scoteseAtlas2021, @straumeGlobal2020, and [VALDES citation]. Each DEM provide paleotopography at resolution of 1&deg; x 1&deg; and 5 myr resolution. We projected the paleoenvironmental reconstructions into an equal-area Behrmann projection of 110km x 110km following methods described in @hagenGen3sis2021. From the projected data we extracted temperature and aridity values across all grid cells that fall within Sahul by manually drawing a polygon that encompassed the continent from 40 mya to the present-day. Because reptiles are sensitive to extremely low and high temperature, we decided to to summarise the reconstructed paleo temperature data into three different data sets: minimum temperature, maximum temperature, and mean temperature.

@@

## Ancestral biome reconstruction and biogeographic history

<!-- ## values and info from 04_BioGeoBEARS_bsm_plots.R -->

All BioGeoBEARS analyses inferred tropical grassland as the ancestral biome of *Anilios*. We identified the DEC+_j_+_x_ model as the most preferred biogeographical history model (delta AICc = -4.47, see Table-supp-BioGeoBEARS). Likelihood ratio test (LRT) indicates that both founder events (+*j* parameter) and distance between biome (+x parameter) together provides statistically significant improvement over the simpler DEC model, DEC+x  and DEC+j (X^2^, DF = 2, D-statistic = 26.7, p-value < 0.001).

We found that _Anilios_ species are mostly found solely or partially in the tropical grassland biome (n = 20) followed by the arid biome (n = 13) and mediterranean biome (n = 10). Closely related species tend to share the same biome (\@ref(fig:fig-biogeobears).

Based on 50 BioGeoBEARS biogeographic stochastic mapping of our best model, while we inferred the tropical grassland biome to be the ancestral biome of Australian _Anilios_ (\@ref(fig:fig-biogeobears)), the ancestor leading to crown *Anilios* is ambiguous. We estimated an average of 71.2 ± 1.3 (mean ± SD) total anagenetic and cladogenetic events occured throughout the phylogeny, out of which the majority were a combination of anagenetic dispersal (31.2 ± 1.3 events) and founder events (6.1 ± 1.5 events) (see supp-summary_counts_bsm_table). 
We found that the largest source of outward dispersal is the tropical grassland biome (16.2 ± 4.9 events; 43.6%) making up almost double the number of outward dispersals from the arid biome (9.7 ± 3.2 events; 26.0%) (\@ref(fig:fig-biogeobears))c-d. Together the tropical grassland biome and the arid biome make up the largest source of diversificttion accounting for 69.6% of outward transitions.

@@

## Diversification rates between arid and non-arid biomes
